---
title: "PCA"
author: "Kate Isaac"
output: html_document
---

# PCA for all respondents

## Load Tidy Data and packages

```{r, message = FALSE}
library(here)
library(tidyverse)

resultsTidy <- readRDS(here("data/wrangled_data/resultsTidy_personas.rds"))
```

## Subset to 

* categorical or a score
* non-free text data 
* isn't related to usertype or persona categorizations

```{r}
pca_data_ar <- resultsTidy %>%
  select(!c(Timestamp,
            CurrentUsageDescription,
            Degrees, #has FurtherSimplifiedDegrees
            ConsortiaAffiliations,
            AllTechPython, #has a Score_ equivalent
            AllTechJupyterNotebooks, #has a Score_ equivalent
            AllTechR, #has a Score_ equivalent
            AllTechRStudio, #has a Score_ equivalent
            AllTechBioconductor, #has a Score_ equivalent
            AllTechGalaxy, #has a Score_ equivalent
            AllTechWorkflows, #has a Score_ equivalent
            AllTechContainers, #has a Score_ equivalent
            AllTechCommandLine, #has a Score_ equivalent
            HumanGenomicExperience, #will use humanGenomicFlag instead
            NonHumanGenomicExperience, #will use nonHumanGenomicFlag instead
            HumanClinicalExperience, #will use clinicalFlag instead
            AnVILDemo, #will use AnVILDemoAwareness and AnVILDemoUse instead
            AnVILSupportForum, #will use forumAwareness and forumUse instead
            ContactWillingness,
            UserType,
            group_col,
            institutionID,
            InstitutionalType,
            clinical_flag, #persona related
            compOnly_flag, #persona related
            education_flag, #persona related
            decision_flag, #persona related
            admin_flag, #persona related
            persona
            )
         )
```

## Combine Potential_ and Returning_ variables to remove NAs when possible and remove Returning_ variable (or variables for returning user specific questions) if there's no equivalent Potential_

```{r}
pca_data_ar <- pca_data_ar %>%
  select(!any_of(c(
    starts_with("ReturningAnVILTech"),
    starts_with("Score_ReturningAnVILTech")
    ))) %>%
  mutate(RankEasyBillingSetup = coalesce(PotentialRankEasyBillingSetup, 
                                         ReturningRankEasyBillingSetup),
         RankFlateRateBilling = coalesce(PotentialRankFlatRateBilling,
                                         ReturningRankFlatRateBilling),
         RankFreeVersion = coalesce(PotentialRankFreeVersion,
                                    ReturningRankFreeVersion),
         RankSupportDocs = coalesce(PotentialRankSupportDocs,
                                    ReturningRankSupportDocs),
         RankToolsData = coalesce(PotentialRankToolsData,
                                  ReturningRankToolsData),
         RankCommunityAdoption = coalesce(PotentialRankCommunityAdoption,
                                          ReturningRankCommunityAdoption)
         ) %>%
  select(!any_of(c(
    starts_with("PotentialRank"),
    starts_with("ReturningRank")
  ))) %>% 
  select(!c(LengthOfUse,
            FeatureMostAppreciated,
            NeededResources,
            RecommendationLikelihood,
            OtherExamplesAnVILTrainingWorkshops
  ))
```


## Turn into vectors 

### Comma separated columns

```{r}
vectorize_comma <- function(pca_data_ar, column_name){
    return(t(
      sapply(strsplit(tolower(pca_data_ar[[column_name]]), ", "), function(x)
        table(factor(x, 
                     levels = unique(unlist(strsplit(tolower(pca_data_ar[[column_name]]), ", ")
                                            )
                                     )
                     )
              )
        )
    ) %>%
      as.data.frame()
  )
}
```

```{r}
#Industry
vectorized_industry <- vectorize_comma(pca_data_ar, "Industry")
  
```

```{r}
#KindOfWork
vectorized_kow <- vectorize_comma(pca_data_ar, "KindOfWork") %>%
  mutate("program_administration" = rowSums(select(., "program administration", "program administration,"), na.rm = TRUE)) %>%
  select(!c("cloud architect", "software development", "program administration", "program administration,")) #remove the ones that got combined and ones provided through "Other" submission
```

```{r}
#TypesOfData
vectorized_datatypes <- vectorize_comma(pca_data_ar, "TypesOfData") %>%
  select(!c("used in training for analysis of genomes (variant calling)", "i don't analyze data on anvil", "i store data in anvil. i don’t analyze it.")) #remove the ones provided through "Other submission"
```

```{r}
#AccessWhichControlledData
vectorized_controlledaccess <- vectorize_comma(pca_data_ar, "AccessWhichControlledData") %>%
  select(!c("gnomad and clinvar", "cancer omics datasets", "tcga", "being able to pull other dbgap data as needed."))
```

```{r}
#WhereAnalysesRun
vectorized_whereanalyses <- vectorize_comma(pca_data_ar, "WhereAnalysesRun") %>%
  mutate(galaxy = rowSums(select(., "galaxy australia", "galaxy (usegalaxy.org)"), na.rm = TRUE),
         hpc = pmin(1, rowSums(select(., "institutional high performance computing cluster (hpc)", "local server"))),
         locally = rowSums(select(., "personal computer (locally)", "personal computer (locally),"))) %>%
  select(!c("sciserver.org", "all of us", "ukbb rap", "institutional high performance computing cluster (hpc)", "local server", "galaxy australia", "galaxy (usegalaxy.org)", "personal computer (locally)", "personal computer (locally),"))
```

```{r}
#RepositoriesDMS
vectorized_dms <- vectorize_comma(pca_data_ar, "RepositoriesDMS") %>%
  select(!c("geo", "encode", "biodata catalyst", "dbgap", "nemo", "sra", "dbgap & ega", "github", "google drive"))
```

```{r}
#FundingSources
vectorized_funding <- vectorize_comma(pca_data_ar, "FundingSources") %>%
  select(!c("nsf"))
```

#### remove columns that were vectorized and add back the columns from the vectorized output

```{r}
pca_data_ar <- pca_data_ar %>%
  select(!c(Industry, 
            KindOfWork, 
            TypesOfData, 
            AccessWhichControlledData, 
            WhereAnalysesRun, 
            RepositoriesDMS, 
            FundingSources)) %>%
  cbind(vectorized_industry, 
        vectorized_kow, 
        vectorized_datatypes, 
        vectorized_controlledaccess, 
        vectorized_whereanalyses, 
        vectorized_dms, 
        vectorized_funding)
```

### Categorical but not factor vectorized yet

```{r}
pca_data_ar <- pca_data_ar %>%
  mutate(across(c(FurtherSimplifiedInstitutionalType, FurtherSimplifiedDegrees, clinicalFlag, humanGenomicFlag, nonHumanGenomicFlag, AnVILDemoAwareness, AnVILDemoUse, forumAwareness, forumUse), ~as.integer(as.factor(.)), .names = "{.col}"))
```

## Confirm the columns are all numeric now

```{r}
pca_data_ar %>% select(where(is.numeric)) %>% ncol() == ncol(pca_data_ar)
```


## Want to keep CurrentUsageDescription/UserType and persona for later visualization

```{r}
annot_data <- resultsTidy %>%
  select(CurrentUsageDescription, UserType, persona)
```


## PCA for all users

```{r}
data.pca <- prcomp(pca_data_ar, center = TRUE, scale. = TRUE)
```

### Elbow plot

```{r}
plot(data.pca, type = "l") #how many PCs to keep before we see the elbow
```

```{r}
summary(data.pca)
```

### Combine annotation data with PCA fit

```{r}
toPlot <- cbind(data.pca$x, annot_data)
```

### PC1 vs PC2 colored by Usertype

```{r}
toPlot %>% ggplot(aes(x = PC1, y = PC2, color = UserType)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

### PC2 vs PC3 colored by Usertype

```{r}
toPlot %>% ggplot(aes(x = PC2, y = PC3, color = UserType)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

### PC1 vs PC3 colored by Usertype

```{r}
toPlot %>% ggplot(aes(x = PC1, y = PC3, color = UserType)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

### PC1 vs PC2 colored by Use description

```{r}
toPlot %>% ggplot(aes(x = PC1, y = PC2, color = CurrentUsageDescription)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

### PC2 vs PC3 colored by Use description

```{r}
toPlot %>% ggplot(aes(x = PC2, y = PC3, color = CurrentUsageDescription)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

### PC1 vs PC3 colored by Use description

```{r}
toPlot %>% ggplot(aes(x = PC1, y = PC3, color = CurrentUsageDescription)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

### PC1 vs PC2 colored by persona

```{r}
toPlot %>% ggplot(aes(x = PC1, y = PC2, color = persona)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

### PC2 vs PC3 colored by persona

```{r}
toPlot %>% ggplot(aes(x = PC2, y = PC3, color = persona)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

### PC1 vs PC3 colored by persona

```{r}
toPlot %>% ggplot(aes(x = PC1, y = PC3, color = persona)) + geom_point() + theme_bw() + theme(legend.position = "bottom")
```

# PCA for just returning users

To do next. Will remove potential user responses and keep the returning user specific data that I had removed for the previous PCA.